---
title: "White Stork Antimicrobial Resistance"
subtitle: "Data Cleaning"
author: "Raquel Francisco and Seth Lattner"
date: "2023-02-24"
output: html_document
---


### Processing script

This QMD file contains processing code for antimicrobial resistance data collected from White stork (*Ciconia ciconia*) in Madrid, Spain.

#### Load Packages/Data

```{r, message=FALSE, warning=FALSE}
#load required packages
library(readr)
library(dplyr)
library(tidyr)
library(skimr) 
library(here) 
```


```{r, message=FALSE, warning=FALSE}
#load in data
stork_AMR_raw <- read_csv(here::here("1 Data Cleaning Script","raw_data","StorkAMRraw_2020_2021.csv"))
```

Let's take a quick look at the data on hand. 

```{r}
dplyr::glimpse(stork_AMR_raw)
skimr::skim(stork_AMR_raw)
```

Just based on this, it looks like there might be a bit of cleaning we'll have to do before we can get much further. Landfill use index (`mean_lui`) data are missing for all observations from the year 2020 due to the COVID-19 pandemic and will thus be excluded from future landfill use analysis. Additionally, there are data missing from the columns `ad`, `age`, `fledged`, and `hatched`. 

No inputs for other variables stand out as being suspicious or blatantly wrong.

#### Data Cleaning

`NA` values in the `fledged` and `hatched` variables are more accurately represented by 0 - no eggs hatched or no eggs fledged. We will replace NA's with 0's to correct this. 

```{r}
d1 <- stork_AMR_raw %>%
  tidyr::replace_na(list(fledged = 0, hatched = 0))
skimr::skim(d1)
```

We would like to add a new column for multi-drug resistance (resistance to ≥3 antibiotic classes) into our dataframe. 

```{r}
#sum across antibiotic classes and create new column when sum ≥3  
d2 <- d1 %>%
  mutate(sum = rowSums(across(c(TET, SULF, AMIN, PHEN, MACR, BLACT, QUIN, POLIMYX)))) %>%
  mutate(MDR = ifelse(sum>2, 1, 0))
glimpse(d2)
```

Standardize main continuous covariates (mean_lui, nsuccess, age)

```{r}
d2$s.nsuccess <- (d2$nsuccess-mean(d2$nsuccess, na.rm = TRUE))/sd(d2$nsuccess, na.rm = TRUE)

#?sd, use na.rm = TRUE to ensure that NAs are removed from calculations
d2$s.lui <- (d2$mean_lui-mean(d2$mean_lui, na.rm = TRUE))/sd(d2$mean_lui, na.rm = TRUE)

d2$s.age <- (d2$age-mean(d2$age, na.rm = TRUE))/sd(d2$age, na.rm = TRUE)

```

Every value below -7 was considered Negative and will be shown as an NA

Quantitative data is reported in log10. Will transform columns to data with the inverse log10 that would allow a more accurate evaluation of AMR gene burden via formula; x=10^y where x is the new gene burden and y is the original gene burden calculated from published formula.

```{r}
# Clean the dataset by setting values below -6.5 to 0 for specified columns
d2 <- d2%>%
  mutate(
    across(
      c(TET..A, TET..B, TET..Y, TET..K, TET..M, TET..Q, TET..S, TET..W, sulI, sulII, str, aadA, catI, catII, ermB, ermF, qnrB, qnrS, blaTEM, mecA, mcr.1),
      ~ ifelse(. < -7, NA, .)
    )
  )

# View the cleaned dataset
glimpse(d2)

# Apply the inverse log10 transformation only to values above 0 and add '_inv10' suffix
d3 <- d2 %>%
  mutate(
    across(
      c(TET..A, TET..B, TET..Y, TET..K, TET..M, TET..Q, TET..S, TET..W, sulI, sulII, str, aadA, catI, catII, ermB, ermF, qnrB, qnrS, blaTEM, mecA, mcr.1),
      ~ ifelse(. > -7, 10^.x, 0),
      .names = "{.col}_inv10"
    )
  ) %>%
  # Calculate Burden_inv10 as the sum of transformed '_inv10' columns
  mutate(Burden_inv10 = rowSums(across(ends_with("_inv10")), na.rm = TRUE)) 

# create sum of log
d4 <- d3 %>%
  mutate(Burden_Log10 = rowSums(across(c(TET..A, TET..B, TET..Y, TET..K, TET..M, TET..Q, TET..S, TET..W, sulI, sulII, str, aadA, catI, catII, ermB, ermF, qnrB, qnrS, blaTEM, mecA, mcr.1)), na.rm = TRUE))

glimpse(d4)
```


We want to make cleaned data file - one with the 2021 data (which includes landfill use) and 2020 which does not.
Let's assign at the end to some final variable, this makes it easier to add further cleaning steps above.

```{r}
#save final
stork_AMR_clean <- d4
```


#### Save Cleaned Data


```{r}
save_data_location2 <- here::here("2 Clean Data","stork_AMR_clean.rds")
saveRDS(stork_AMR_clean, file = save_data_location2)

# Save as a CSV file for cross-platform compatibility
save_data_location3 <- here::here("2 Clean Data", "stork_AMR_clean.csv")
write.csv(stork_AMR_clean, file = save_data_location3, row.names = FALSE)
```