---
title: "White Stork Antimicrobial Resistance"
subtitle: "Data Cleaning"
author: "Raquel Francisco and Seth Lattner"
date: "2023-02-24"
output: html_document
---


### Processing script

This QMD file contains processing code for antimicrobial resistance data collected from White stork (*Ciconia ciconia*) in Madrid, Spain.

#### Load Packages/Data

```{r, message=FALSE, warning=FALSE}
#load required packages
library(readr)
library(dplyr)
library(tidyr)
library(skimr) 
library(here) 
```


```{r, message=FALSE, warning=FALSE}
#load in data
stork_AMR_raw <- read_csv(here::here("1 Data Cleaning Script","raw_data","StorkAMRraw_2020_2021.csv"))
```

Let's take a quick look at the data on hand. 

```{r}
dplyr::glimpse(stork_AMR_raw)
skimr::skim(stork_AMR_raw)
```

Just based on this, it looks like there might be a bit of cleaning we'll have to do before we can get much further. Landfill use index (`mean_lui`) data are missing for all observations from the year 2020 due to the COVID-19 pandemic and will thus be excluded from future landfill use analysis. Additionally, there are data missing from the columns `ad`, `age`, `fledged`, and `hatched`. 

No inputs for other variables stand out as being suspicious or blatantly wrong.

#### Data Cleaning

`NA` values in the `fledged` and `hatched` variables are more accurately represented by 0 - no eggs hatched or no eggs fledged. We will replace NA's with 0's to correct this. 

```{r}
d1 <- stork_AMR_raw %>%
  tidyr::replace_na(list(fledged = 0, hatched = 0))
skimr::skim(d1)
```

We would like to add a new column for multi-drug resistance (resistance to ≥3 antibiotic classes) into our dataframe. 

```{r}
#sum across antibiotic classes and create new column when sum ≥3  
d2 <- d1 %>%
  mutate(sum = rowSums(across(c(TET, SULF, AMIN, PHEN, MACR, BLACT, QUIN, POLIMYX)))) %>%
  mutate(MDR = ifelse(sum>2, 1, 0))
glimpse(d2)
```

Standardize main continuous covariates (mean_lui, nsuccess, age)

```{r}
d2$s.nsuccess <- (d2$nsuccess-mean(d2$nsuccess, na.rm = TRUE))/sd(d2$nsuccess, na.rm = TRUE)

#?sd, use na.rm = TRUE to ensure that NAs are removed from calculations
d2$s.lui <- (d2$mean_lui-mean(d2$mean_lui, na.rm = TRUE))/sd(d2$mean_lui, na.rm = TRUE)

d2$s.age <- (d2$age-mean(d2$age, na.rm = TRUE))/sd(d2$age, na.rm = TRUE)

```

Quantitative data is reported in log10. Will transform columns to data with the inverse log10 that would allow a more accurate evaluation of AMR gene burden via formula; x=10^y where x is the new gene burden and y is the original gene burden calculated from published formula.

```{r}
d2$TET..A <- 10^(d2$TET..A)
d2$TET..B <- 10^(d2$TET..B)
d2$TET..Y <- 10^(d2$TET..Y)
d2$TET..K <- 10^(d2$TET..K)
d2$TET..M <- 10^(d2$TET..M)
d2$TET..Q <- 10^(d2$TET..Q)
d2$TET..S <- 10^(d2$TET..S)
d2$TET..W <- 10^(d2$TET..W)
d2$sulI <- 10^(d2$sulI)
d2$sulII <- 10^(d2$sulII)
d2$str <- 10^(d2$str)
d2$aadA <- 10^(d2$aadA)
d2$catI <- 10^(d2$catI)
d2$catII <- 10^(d2$catII)
d2$ermB <- 10^(d2$ermB)
d2$ermF <- 10^(d2$ermF)
d2$qnrB <- 10^(d2$qnrB)
d2$qnrS <- 10^(d2$qnrS)
d2$blaTEM <- 10^(d2$blaTEM)
d2$mecA <- 10^(d2$mecA)
d2$mcr.1 <- 10^(d2$mcr.1)

d3 <- d2 %>%
  mutate(Burden = rowSums(across(c(TET..A, TET..B, TET..Y, TET..K, TET..M, TET..Q, TET..S, TET..W, sulI, sulII, str, aadA, catI, catII, ermB, ermF, qnrB, qnrS, blaTEM, mecA, mcr.1))))

glimpse(d3)
```


We want to make two cleaned data files - one with only the 2021 data (which includes landfill use) and one with both 2020 and 2021 data.

```{r}
#filter out 2020 data
d4 <- d3 %>% filter(yr !=2020)
head(d4)
```


Let's assign at the end to some final variable, this makes it easier to add further cleaning steps above.

```{r}
#save 2021 data only
stork_AMR_2021 <- d4

#save for both years
stork_AMR_clean <- d3
```


#### Save Cleaned Data


```{r}
save_data_location <- here::here("2 Clean Data","stork_AMR_2021.rds")
saveRDS(stork_AMR_2021, file = save_data_location)

save_data_location2 <- here::here("2 Clean Data","stork_AMR_clean.rds")
saveRDS(stork_AMR_clean, file = save_data_location2)
```