---
title: "White Stork Antimicrobial Resistance"
subtitle: "Exploratory Analysis"
author: "Raquel Francisco and Seth Lattner"
date: "2022-02-24"
output: html_document
---

This QMD file contains exploratory analysis for antimicrobial resistance data collected from White Stork (*Ciconia ciconia*) in Madrid, Spain.
  

#### Load Data/Packages

```{r, warning=FALSE, message=FALSE}
#load needed packages
library(here) #calling
library(dplyr) #cleaning
library(skimr)
library(ggplot2) #plotting and visualization
library(ggpubr) #ggscatter to visualize correlations
library(naniar)#missing values graphics and plots
library(UpSetR)#missing values graphics and plots
library(corrplot) #evaluate if covariates are correlated
library(writexl)
library(reshape2)
library(tidyverse)
library(tidyr)
library(knitr)
library(nortest)
```


```{r}
#load all data
stork_AMR_clean <- readRDS(here::here("2 Clean Data","stork_AMR_clean.rds"))

# Subset the dataset for the year 2021
stork_AMR_2021 <- stork_AMR_clean %>%
  filter(yr == 2021)

stork_AMR_2020 <- stork_AMR_clean %>%
  filter(yr == 2020)

```

#### Quality Control

```{r}
stork_AMR_clean %>%
  as_shadow_upset() %>%
  upset()

gg_miss_fct(x = stork_AMR_clean, fct = yr)
gg_miss_fct(x = stork_AMR_clean, fct = samp)
```
Quite alot of missing values for 2020. Will likely only be able to do summary stats on 2020 because no LUI data was collected that year (COVID-19 limitations)

#### Data subset for 2021

```{r}
stork_AMR_2021 %>%
  as_shadow_upset() %>%
  upset()

gg_miss_fct(x = stork_AMR_2021, fct = yr)
gg_miss_fct(x = stork_AMR_2021, fct = samp)
```

#### Correlation Check of Covariates

```{r}
#define the variables we want to include in the correlation matrix
variables <- stork_AMR_clean[ , c("s.age", "s.lui", "samp", "s.nsuccess")]

#create correlation matrix
cor(variables, use = "complete.obs") #ignore NAs
CorrPlot <- corrplot(cor(variables, use = "complete.obs"))
CorrPlot

Z <- data.frame(s.age = stork_AMR_clean$s.age, 
                s.lui = stork_AMR_clean$s.lui, 
                s.nsuccess = stork_AMR_clean$s.nsuccess,
                samp = stork_AMR_clean$samp)
pairs(Z, panel = panel.smooth, labels = c("age", "mean_lui", "nsuccess", "samp"))

#More than 0.7 usually is unacceptable
#More than 0.5 you need to be careful; NOTHING over 0.5

##BASIC SPEARMAN
cor.test( stork_AMR_clean$s.age, stork_AMR_clean$s.lui,
         method = "spearman")
cor.test( stork_AMR_clean$s.age, stork_AMR_clean$s.nsuccess,
          method = "spearman")
cor.test( stork_AMR_clean$s.age, stork_AMR_clean$samp,
          method = "spearman")
cor.test( stork_AMR_clean$s.nsuccess, stork_AMR_clean$s.lui,
          method = "spearman")
cor.test( stork_AMR_clean$s.nsuccess, stork_AMR_clean$samp,
          method = "spearman")
cor.test( stork_AMR_clean$s.lui, stork_AMR_clean$samp,
          method = "spearman")
cor.test( stork_AMR_clean$Burden_inv10, stork_AMR_clean$MDR,
          method = "spearman")
cor.test( stork_AMR_clean$Burden_Log10, stork_AMR_clean$MDR,
          method = "spearman")
cor.test( stork_AMR_clean$sum, stork_AMR_clean$MDR,
          method = "spearman") #Correlated!

##Visual Plots with: regression line and confidence interval

#Age + LUI
ggscatter(stork_AMR_clean, x = 'age', y = 's.lui',
          add = "reg.line", conf.int = TRUE,    
          add.params = list(fill = "lightgray"),
          ggtheme = theme_minimal()
)+
  stat_cor(method = "spearman", p.accuracy = 0.001, r.accuracy = 0.01,
           label.y = 0.5) 

#Nsuccess + LUI
ggscatter(stork_AMR_clean, x = 's.nsuccess', y = 's.lui',
          add = "reg.line", conf.int = TRUE,    
          add.params = list(fill = "lightgray"),
          ggtheme = theme_minimal()
)+
  stat_cor(method = "spearman", p.accuracy = 0.001, r.accuracy = 0.01,
           label.y = 0.6)

#Nsuccess + Age
ggscatter(stork_AMR_clean, x = 's.nsuccess', y = 'age',
          add = "reg.line", conf.int = TRUE,    
          add.params = list(fill = "lightgray"),
          ggtheme = theme_minimal()
)+
  stat_cor(method = "spearman", p.accuracy = 0.001, r.accuracy = 0.01,
           label.y = 20)

#Variables of interest not correlated; they are independent = Move on to analysis
```


#### Data Exploration and visualization

Will explore both 2020 and 2021 data, but will not model with 2020 as no LUI data was collected this year due to COVID-19 resrictions.

Quick Sample summary stats

```{r}
# Total number of samples
total_samples <- nrow(stork_AMR_clean)
cat("Total number of samples:", total_samples, "\n")

# Number of samples in each year
samples_per_year <- stork_AMR_clean %>%
  group_by(yr) %>%
  summarise(count = n())
cat("Samples per year:\n")
print(samples_per_year)

# Breakdown of samples by year and sampling period (samp)
samples_by_year_and_samp <- stork_AMR_clean %>%
  group_by(yr, samp) %>%
  summarise(count = n())
cat("Samples by year and sampling period:\n")
print(samples_by_year_and_samp)

```

Samples per sampling period per year
```{r}
# Create a summary table for 2020
summary_2020 <- stork_AMR_2020 %>%
  group_by(samp) %>%
  summarise(
    nests = paste(unique(nes), collapse = ", "), # List all nests
    total_samples = n() # Count total samples
  ) %>%
  mutate(year = 2020)

# Create a summary table for 2021
summary_2021 <- stork_AMR_2021 %>%
  group_by(samp) %>%
  summarise(
    nests = paste(unique(nes), collapse = ", "), # List all nests
    total_samples = n() # Count total samples
  ) %>%
  mutate(year = 2021)

# Combine both years into a single table
summary_table <- bind_rows(summary_2020, summary_2021) %>%
  arrange(year, samp)

# Display the table
kable(summary_table, caption = "Summary of Nests and Samples by Sampling Period for 2020 and 2021")

# Optionally, save the table to a CSV file
#write.csv(summary_table, "nest_sampling_summary.csv", row.names = FALSE)


# Reshape data for 2020
nests_2020 <- stork_AMR_2020 %>%
  group_by(nes) %>%
  summarise(sampling_periods = paste(sort(unique(samp)), collapse = ", ")) %>%
  mutate(year = 2020)

# Reshape data for 2021
nests_2021 <- stork_AMR_2021 %>%
  group_by(nes) %>%
  summarise(sampling_periods = paste(sort(unique(samp)), collapse = ", ")) %>%
  mutate(year = 2021)

# Combine both years for an overview, if needed
combined_nests <- bind_rows(nests_2020, nests_2021)

# Display the tables for 2020 and 2021
kable(nests_2020, caption = "Nests and Sampling Periods for 2020")
kable(nests_2021, caption = "Nests and Sampling Periods for 2021")

# Save each year's table to a CSV file
write.csv(nests_2020, here("4 Results", "Tables", "nests_2020.csv"), row.names = FALSE)
write.csv(nests_2021, here("4 Results", "Tables", "nests_2021.csv"), row.names = FALSE)
write.csv(combined_nests, here("4 Results", "Tables", "nests_20202021.csv"), row.names = FALSE)
```



First we want to make a table to take a quick look at a few variables that will be important later on.

```{r}
table1 <- stork_AMR_clean %>%
  select(id, age, nsuccess, mean_lui, Burden_inv10, MDR) %>%
  knitr::kable()
table1
```

Plotting total Burden outcomes from both years.

```{r}
stork_AMR_clean %>%
  ggplot(aes(as.factor(yr), Burden_inv10))+
  geom_boxplot()
```
Plotting total Burden outcomes across all sampling periods.
```{r}
stork_AMR_clean %>%
  ggplot(aes(as.factor(samp), Burden_inv10))+
  geom_boxplot()

pBurden <- stork_AMR_clean %>%
  ggplot(aes(as.factor(samp), Burden_inv10)) +
  geom_boxplot() +
  geom_jitter(width = 0.2, alpha = 0.5) +  # Add jittered points with slight transparency
  ylab('Total Resistance Gene Burden (Inverse Log10)') +
  xlab('') +
 scale_x_discrete(labels = c("1" = "Sampling Period 1", "2" = "Sampling Period Period 2", "3" = "Sampling Period 3"))  

# Display the plot
pBurden

# Save the plot as TIFF using ggsave
ggsave(filename = here('4 Results', 'Figures', 'Burden_Sample.tiff'), 
       plot = pBurden, device = "tiff", width = 10, height = 6, dpi = 300)
```
Sampling period and MDR

```{r}
summary_stats <- stork_AMR_clean %>%
  group_by(samp) %>%
  summarise(
    mean_MDR = mean(MDR),  # Proportion of MDR = 1
    se_MDR = sqrt(mean_MDR * (1 - mean_MDR) / n())  # Standard error of the proportion
  )

# Plot with box plot and error bars
pMDR <- stork_AMR_clean %>%
  ggplot(aes(x = as.factor(samp), y = MDR)) +
  geom_boxplot(alpha = 0.6, outlier.shape = NA) +  # Single box plot per samp
  geom_jitter(position = position_jitter(width = 0.2, height = 0), 
              alpha = 0.5, color = "black") +  # Add raw data points
  geom_errorbar(data = summary_stats,  # Error bars on top
                aes(x = as.factor(samp), 
                    ymin = mean_MDR - se_MDR, 
                    ymax = mean_MDR + se_MDR), 
                width = 0.2, color = "black", size = 0.8, inherit.aes = FALSE) + 
  geom_point(data = summary_stats,  # Mean points on top
             aes(x = as.factor(samp), y = mean_MDR), 
             color = "black", size = 3, shape = 21, fill = "white", inherit.aes = FALSE) +
  ylab('Multi-Drug Resistance') +
  xlab('') +
  scale_x_discrete(labels = c("1" = "Sampling Period 1", 
                              "2" = "Sampling Period 2", 
                              "3" = "Sampling Period 3")) 


# Display the plot
pMDR

# Save the plot as TIFF using ggsave
ggsave(filename = here('4 Results', 'Figures', 'MDR_Sample.tiff'), 
       plot = pMDR, device = "tiff", width = 10, height = 6, dpi = 300)

```



Total Number of AMR Gene Classes over both years.

```{r}
pSB <- stork_AMR_clean %>%
  ggplot(aes(as.factor(yr), sum))+
  geom_boxplot() +
  ylab('Breeding Season Sampling Period') +
  xlab ('Antimicrobial Gene Resistance Burden')

# Save the plot as TIFF using ggsave
ggsave(filename = here('4 Results', 'Figures', 'Prelim_BurdenvsSamp.tiff'), 
       plot = pSB, device = "tiff", width = 10, height = 6, dpi = 300)
```

Sample vs Total Number of AMR Gene Classes
```{r}
stork_AMR_clean %>%
  ggplot(aes(as.factor(samp), sum))+
  geom_boxplot()
```

Plot Gene Burden vs age of bird 

```{r}
stork_AMR_clean %>%
  ggplot(aes(age, Burden_Log10))+
  geom_point()
```

Plot MDR vs age of bird 

```{r}
stork_AMR_clean %>%
  ggplot(aes(as.factor(MDR), age))+
  geom_boxplot()
```

Plot Gene Burden vs age of bird 

```{r}
stork_AMR_clean %>%
  ggplot(aes(nsuccess, Burden_Log10))+
  geom_point()
```

Plot MDR vs nest success

```{r}
stork_AMR_clean %>%
  ggplot(aes(as.factor(MDR), nsuccess))+
  geom_boxplot()
```

Now we will make a table filtering out 2020.

```{r}
table2 <- stork_AMR_2021 %>%
  select(id, age, nsuccess, mean_lui, Burden_Log10, MDR) %>%
  knitr::kable()
table2
```


Plotting mean landfill use index vs nest success

```{r, warning=FALSE, message=FALSE}
stork_AMR_2021 %>%
  ggplot(aes(mean_lui, nsuccess))+
  geom_point()+
  geom_smooth(method = "lm")
```

Plotting mean landfill use index vs nest success, grouped by MDR outcome.

```{r, message=FALSE, warning=FALSE}
pLN <- stork_AMR_clean%>%
  ggplot(aes(mean_lui, nsuccess, color=as.factor(MDR)))+
  geom_point()+
  geom_smooth(method = "lm")+
  ylab('Nest Success %') +
  xlab ('Landfill Use') +
  scale_x_continuous(breaks = c(0, 0.1, 0.2, 0.3, 0.4),   # Define specific x-axis breaks
                     labels = c("None (0)", "Mild (0.1)", "Moderate (0.2)", "High (0.3)", "Very High (0.4)"))  +
  scale_color_manual(name = "Multi-Drug Resistance",
                     labels = c("0" = "Absent", "1" = "Present"),  # Custom labels for the legend
                     values = c("0" = "darkred", "1" = "forestgreen"))  # Custom colors for each level


pLN

# Save the plot as a TIFF file
ggsave(filename = here("4 Results", "Figures", "RAWluiVSns.tiff"), 
       plot = pLN, 
       device = "tiff", 
       width = 10, 
       height = 6, 
       dpi = 300)
```

Plotting mean landfill use index vs nest success, grouped by MDR outcome.

```{r, message=FALSE, warning=FALSE}
stork_AMR_2021%>%
  ggplot(aes(mean_lui, nsuccess, color=Burden_inv10))+
  geom_point()+
  geom_smooth(method = "lm")
```

Plotting MDR vs mean landfill use index

```{r, warning=FALSE}
pLM <- stork_AMR_2021 %>%
  ggplot(aes(as.factor(MDR), mean_lui))+
  geom_boxplot()+
  geom_jitter(width = 0.2, alpha = 0.5) +  # Add jittered points with transparency
  ylab('Landfill Use') +
  xlab ('Multi-Drug Resistance')+
  scale_x_discrete(labels = c("0" = "Absent", "1" = "Present"))  +  # Map 0 to "No" and 1 to "Yes"
  scale_y_continuous(breaks = c(0, 0.1, 0.2, 0.3, 0.4),   # Define specific y-axis breaks
                     labels = c("None (0)", "Mild (0.1)", "Moderate (0.2)", "High (0.3)", "Very High (0.4)"))  # Custom labels for y-axis

pLM

#save as tiff
tiff(file = here('4 Results', 'Figures', 'Prelim_LUIvsMD.tiff'))
pLM
dev.off() 
```

Plotting Burden vs mean landfill use index

```{r, warning=FALSE}
stork_AMR_2021 %>%
  ggplot(aes(mean_lui, Burden_Log10))+
  geom_point()+
  geom_smooth(method = "lm")
```

There seems to be a slight decrease in the presence of MDR and increase in the burden of AMR genes as landfill use index increases. Otherwise, there are not many apparent trends in the data off the bat. This exploration will help us as we move forward into a more robust statistical analysis.

Lets go back and just visulaize raw data.
Gene Heat Map
Log10 tranform data first
```{r}
# Define gene columns for log transformation
gene_columns <- c("TET..A", "TET..B", "TET..Y", "TET..K", "TET..M", 
                  "TET..Q", "TET..S", "TET..W", "sulI", "sulII", "str", 
                  "aadA", "catI", "catII", "ermB", "ermF", "mecA", "mcr.1", 
                  "qnrB", "blaTEM", "qnrS")


# Convert data to long format using the specified gene columns
data_long <- stork_AMR_clean %>%
  pivot_longer(
    cols = all_of(gene_columns),
    names_to = "Gene",
    values_to = "Abundance"
  )

# Separate data by year
data_2020 <- data_long %>% filter(yr == 2020)
data_2021 <- data_long %>% filter(yr == 2021)

create_and_save_heatmap <- function(data, year, filepath) {
  heatmap <- ggplot(data, aes(x = nes, y = Gene, fill = Abundance)) +
    geom_tile(na.rm = FALSE) + # Ensures NA values are kept
    facet_wrap(~samp, scales = "free", nrow = 1) +
    labs(title = paste("Gene Abundance across Sampling Periods for", year),
         x = "Nest ID", y = "Resistance Gene") +
    # Use a two-color gradient with NA values as white
    scale_fill_gradient(
      low = "dodgerblue2", 
      high = "red", 
      na.value = "white", # Set NA values to white
      name = "Abundance"
    ) +
    theme_minimal() +
    scale_x_discrete(guide = guide_axis(n.dodge = 2)) +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 8))
  
  # Save the plot as a TIFF file
  ggsave(filename = filepath, plot = heatmap, device = "tiff", width = 10, height = 6, dpi = 300)
}

# Save heatmaps for each year
create_and_save_heatmap(data_2020, "2020", here('4 Results', 'Figures', 'Prelim_LUIvsMD_2020.tiff'))
create_and_save_heatmap(data_2021, "2021", here('4 Results', 'Figures', 'Prelim_LUIvsMD_2021.tiff'))
```

Lets just quickly review normality of the Burden Variables/Distribution

Must use GLMM for both
```{r}
# Subset relevant variables
burden_data <- stork_AMR_clean %>% select(Burden_Log10, Burden_inv10)

# Perform Shapiro-Wilk test for normality
shapiro_log10 <- shapiro.test(burden_data$Burden_Log10)
shapiro_inv10 <- shapiro.test(burden_data$Burden_inv10)

# Perform Kolmogorov-Smirnov test for normality
ks_log10 <- ks.test(burden_data$Burden_Log10, "pnorm", mean(burden_data$Burden_Log10, na.rm=TRUE), sd(burden_data$Burden_Log10, na.rm=TRUE))
ks_inv10 <- ks.test(burden_data$Burden_inv10, "pnorm", mean(burden_data$Burden_inv10, na.rm=TRUE), sd(burden_data$Burden_inv10, na.rm=TRUE))

# Print test results
cat("Shapiro-Wilk Test Results:\n")
print(shapiro_log10)
print(shapiro_inv10)

cat("\nKolmogorov-Smirnov Test Results:\n")
print(ks_log10)
print(ks_inv10)

# Generate histograms for visualization
p1 <- ggplot(stork_AMR_clean, aes(x = Burden_Log10)) +
  geom_histogram(aes(y = ..density..), bins = 30, fill = "blue", alpha = 0.5) +
  geom_density(color = "red", size = 1) +
  ggtitle("Histogram and Density Plot: Burden_Log10") +
  theme_minimal()

p2 <- ggplot(stork_AMR_clean, aes(x = Burden_inv10)) +
  geom_histogram(aes(y = ..density..), bins = 30, fill = "blue", alpha = 0.5) +
  geom_density(color = "red", size = 1) +
  ggtitle("Histogram and Density Plot: Burden_inv10") +
  theme_minimal()

# Generate Q-Q plots for normality assessment
p3 <- ggplot(stork_AMR_clean, aes(sample = Burden_Log10)) +
  stat_qq() +
  stat_qq_line() +
  ggtitle("Q-Q Plot: Burden_Log10") +
  theme_minimal()

p4 <- ggplot(stork_AMR_clean, aes(sample = Burden_inv10)) +
  stat_qq() +
  stat_qq_line() +
  ggtitle("Q-Q Plot: Burden_inv10") +
  theme_minimal()

# Display the plots
print(p1)
print(p2)
print(p3)
print(p4)

```

