---
title: "StatiscalAnalysis_ModelFitting"
format: html
editor: visual
---

# Ciconia ciconia analysis script

#### this script loads the processed, cleaned data, does a simple analysis and saves the results to the results folder

```{r, warning=FALSE, message=FALSE}
#load needed packages. make sure they are installed.
library(tidyverse)
library(ggplot2) #for plotting
library(ggeffects) #to explore plots  
library(broom) #for cleaning up output from lm()
library(broom.mixed) #for cleaning up output from lmer()
library(here) #for data loading/saving
library(multilevelmod) #mixed effect models
library(parsnip)
library(lme4)
library(sjPlot) #to help plot models
library(performance) #evaluate model fit and performance
library(see)
library(cAIC4)
library(flextable)
library(officer)
library(dplyr) #cleaning
library(patchwork)
library(tidymodels)

#load all data
stork_AMR_clean <- readRDS(here::here("2 Clean Data","stork_AMR_clean.rds"))

# Subset the dataset for the year 2021
data <- stork_AMR_clean %>%
  filter(yr == 2021)


```

## Data fitting/statistical analysis

Data of interest: Landfill use index (LUI) and multi-drug resistence (MDR) throughout the different sampling periods during the stork breeding season.

We'll start by fitting two models. We will used mixed-effects models to control for repeated sampling of the same nest. The design is unbalanced, thus we cannot use a repeated measures anova.

We'll start with a single factor repeated measures evaluation the effect of LUI on blaTEM (betalactam) resistence gene burden.

```{r}
facet_plot <- data %>% 
  ggplot(aes(x = samp, y = Burden_inv10)) + 
  geom_point() + 
  geom_line() +
  facet_wrap(~ nes) 
facet_plot

# Use ggsave to save the plot
ggsave(filename = here('4 Results', 'Figures', 'facet_plot.png'), plot = facet_plot, width = 10, height = 6, dpi = 300)
```

No obvious trends...

Build models using glmer

```{r}
#MDR
glmer_spec <- 
  logistic_reg() %>% 
  set_engine("glmer")
```

Multi-drug resistance (0,1 data) vs Covariates
Check the varience of the fixed effect (Nest = Nes)

```{r}
# Define models (GLMMs with nest (nes) as a random effect)
glmer_fit_null <- glmer(as.factor(MDR) ~ (1 | nes), 
                        data = data, 
                        family = binomial(link = "logit"),
                        control = glmerControl(optimizer = "bobyqa"))

glmer_fit_global <- glmer(as.factor(MDR) ~ s.lui + s.age + samp + s.nsuccess + (1 | nes), 
                          data = data, 
                          family = binomial(link = "logit"),
                          control = glmerControl(optimizer = "bobyqa"))

# Univariate Models
glmer_fit_1 <- glmer(as.factor(MDR) ~ s.lui + (1 | nes), data = data, family = binomial(link = "logit"))
glmer_fit_2 <- glmer(as.factor(MDR) ~ s.age + (1 | nes), data = data, family = binomial(link = "logit"))
glmer_fit_3 <- glmer(as.factor(MDR) ~ samp + (1 | nes), data = data, family = binomial(link = "logit"))
glmer_fit_4 <- glmer(as.factor(MDR) ~ s.nsuccess + (1 | nes), data = data, family = binomial(link = "logit"))

# Multivariate Models
glmer_fit_5 <- glmer(as.factor(MDR) ~ s.lui + samp + (1 | nes), data = data, family = binomial(link = "logit"))
glmer_fit_6 <- glmer(as.factor(MDR) ~ s.lui + s.nsuccess + (1 | nes), data = data, family = binomial(link = "logit"))
glmer_fit_7 <- glmer(as.factor(MDR) ~ samp + s.nsuccess + (1 | nes), data = data, family = binomial(link = "logit"))
glmer_fit_8 <- glmer(as.factor(MDR) ~ s.lui * s.age + samp + s.nsuccess + (1 | nes), data = data, family = binomial(link = "logit"))
glmer_fit_9 <- glmer(as.factor(MDR) ~ s.lui + s.age * samp + s.nsuccess + (1 | nes), data = data, family = binomial(link = "logit"))
glmer_fit_10 <- glmer(as.factor(MDR) ~ s.lui + s.age + samp * s.nsuccess + (1 | nes), data = data, family = binomial(link = "logit"))
glmer_fit_11 <- glmer(as.factor(MDR) ~ s.lui + s.age + (1 | nes), data = data, family = binomial(link = "logit"))
glmer_fit_12 <- glmer(as.factor(MDR) ~ s.lui + s.age + s.nsuccess + (1 | nes), data = data, family = binomial(link = "logit"))

## **Step 2: Check Variance of `nes` for Each Model**

# Create a list of all GLMM models
glmer_models <- list(
  glmer_fit_null, glmer_fit_global, glmer_fit_1, glmer_fit_2, glmer_fit_3, 
  glmer_fit_4, glmer_fit_5, glmer_fit_6, glmer_fit_7, glmer_fit_8, 
  glmer_fit_9, glmer_fit_10, glmer_fit_11, glmer_fit_12
)

# Function to check variance of random effect
check_variance <- function(model, model_name) {
  variance <- summary(model)$varcor
  singular <- isSingular(model, tol = 1e-4)
  
  return(data.frame(
    Model = model_name,
    Variance = ifelse(length(variance) > 0, as.numeric(variance[[1]][1]), NA),
    Singular = singular
  ))
}

# Check variance for all models
variance_results <- do.call(rbind, 
                            lapply(seq_along(glmer_models), function(i) {
                              check_variance(glmer_models[[i]], paste0("Model_", i))
                            }))

# Print results
print(variance_results)

```

Keep NES as a random effect since the varience is high!

Use AIC to see the best model; None look great...

```{r}
glmer_indices <- compare_performance(glmer_fit_null,glmer_fit_global, glmer_fit_1, glmer_fit_2, glmer_fit_3, glmer_fit_4, glmer_fit_5, glmer_fit_6, glmer_fit_7, glmer_fit_8, glmer_fit_9, glmer_fit_10, glmer_fit_11, glmer_fit_12)

print(glmer_indices)

# Create and format the AIC table
AIC_table1 <- compare_performance(glmer_fit_null, glmer_fit_global, glmer_fit_1, glmer_fit_2, glmer_fit_3, 
                                  glmer_fit_4, glmer_fit_5, glmer_fit_6, glmer_fit_7, glmer_fit_8, 
                                  glmer_fit_9, glmer_fit_10, glmer_fit_11, glmer_fit_12, 
                                  rank = TRUE, metrics = "AICc")

# Confirm all models are in the table
print(AIC_table1)

# Convert to a flextable and format with 3 decimal places
AIC_table_flextable <- AIC_table1 %>%
  as_flextable() %>%
  set_header_labels(.,
                    Name = "Model",
                    AICc_wt = "AICc Weight",
                    Performance_Score = "Performance Score") %>%
  colformat_num(j = "AICc_wt", digits = 3) %>%  # Format AICc Weight column with 3 decimal places
  colformat_num(j = "Performance_Score", digits = 3) %>%  # Format Performance Score with 3 decimals if applicable
  theme_box() %>%  # Add a simple box theme
  autofit()  # Adjust column widths automatically

# Save as a Word document
doc <- read_docx() %>% 
  body_add_par("Table: AICc Model Comparison", style = "heading 1") %>%  # Add a title if needed
  body_add_flextable(AIC_table_flextable)

# Define the path and save
print(doc, target = here("4 Results", "Tables", "ModelAIC1.docx"))

#Tiff 
tiff(file = here('4 Results', 'Figures', 'glmer_indices.tiff'))
glmer_indices

```

Summarize and graph best model (lowwest AIC)

```{r}
# Generate the check_model plot
glmer_check <- check_model(glmer_fit_2)
glmer_check

# Access individual plots and save each
ggsave(filename = here('4 Results', 'Figures', 'glmer_check_qqplot.tiff'), 
       plot = glmer_check$plots[[1]], width = 8, height = 6, dpi = 300)

ggsave(filename = here('4 Results', 'Figures', 'glmer_check_residuals.tiff'), 
       plot = glmer_check$plots[[2]], width = 8, height = 6, dpi = 300)

# place results from fit into a data frame with the tidy function

model1 <- broom::tidy(glmer_fit_2)
print(model1)

# Convert to flextable and format with 3 decimal places
model1_flextable <- model1 %>%
  as_flextable() %>%
  set_header_labels(
    term = "Term",
    estimate = "Estimate",
    std.error = "Standard Error",
    statistic = "Z-value",
    p.value = "P-value"
  ) %>%
  colformat_num(j = c("estimate", "std.error", "statistic", "p.value"), digits = 3) %>%  # Format columns with 3 decimal places
  theme_box() %>%
  autofit()

# Save as a Word document
doc <- read_docx() %>% 
  body_add_par("Table: Model Summary", style = "heading 1") %>%  # Add a title
  body_add_flextable(model1_flextable)

# Define the path and save the Word document
print(doc, target = here("4 Results", "Tables", "ModelSummary.docx"))

# save fit results table  
model1 %>%
  saveRDS(here("4 Results", "Models", "GLMERmodel1.rds"))

MLp <- plot_model(glmer_fit_2,terms = 's.age', type="pred",
           ci.lvl = 0.95,  # Ensures 95% confidence interval
           show.data = TRUE, show.values = TRUE, show.p = TRUE,
           title = "", show.legend = TRUE, line.size=1,
           axis.title = c("Standardized Age",'Multi-Drug Resistence Status'))  +
           scale_y_continuous(breaks = c(0, 1), labels = c("Absent", "Present")) + 
           scale_x_continuous(breaks = c(-1, 0, 1, 2))

MLp 

  while (!is.null(dev.list()))  dev.off()
#Tiff 
tiff(file = here('4 Results', 'Figures', 'AGEvsMDR_LME.tiff'))
MLp
```

AMR Gene Burden vs All Since Burden_Log10 is continuous but not normally distributed, use glmer() from lme4 with a Gamma (log-link) family.

Try to use Burden_inv10 since no NEGs

Still all singular, consider removing NES. When tested random effect for nes (nest ID) has an extremely small variance (at or near 0), which suggests that it is contributing almost no meaningful variation to the model.

```{r}
# Null model (random intercept only)
glmer_fit2_null <- glmer(Burden_inv10 ~ (1 | nes), 
                         data = data, 
                         family = Gamma(link = "log"),
                         control = glmerControl(optimizer = "bobyqa"))

# Global model with all predictors
glmer_fit2_global <- glmer(Burden_inv10 ~ s.lui + s.age + samp + s.nsuccess + (1 | nes), 
                           data = data, 
                           family = Gamma(link = "log"),
                           control = glmerControl(optimizer = "bobyqa"))

# Univariate models
glmer_fit2_1 <- glmer(Burden_inv10 ~ s.lui + (1 | nes), 
                      data = data, 
                      family = Gamma(link = "log"),
                      control = glmerControl(optimizer = "bobyqa"))

glmer_fit2_2 <- glmer(Burden_inv10 ~ s.age + (1 | nes), 
                      data = data, 
                      family = Gamma(link = "log"),
                      control = glmerControl(optimizer = "bobyqa"))

glmer_fit2_3 <- glmer(Burden_inv10 ~ samp + (1 | nes), 
                      data = data, 
                      family = Gamma(link = "log"),
                      control = glmerControl(optimizer = "bobyqa"))

glmer_fit2_4 <- glmer(Burden_inv10 ~ s.nsuccess + (1 | nes), 
                      data = data, 
                      family = Gamma(link = "log"),
                      control = glmerControl(optimizer = "bobyqa"))

# Multivariate models
glmer_fit2_5 <- glmer(Burden_inv10 ~ s.lui + samp + (1 | nes), 
                      data = data, 
                      family = Gamma(link = "log"),
                      control = glmerControl(optimizer = "bobyqa"))

glmer_fit2_6 <- glmer(Burden_inv10 ~ s.lui + s.nsuccess + (1 | nes), 
                      data = data, 
                      family = Gamma(link = "log"),
                      control = glmerControl(optimizer = "bobyqa"))

glmer_fit2_7 <- glmer(Burden_inv10 ~ samp + s.nsuccess + (1 | nes), 
                      data = data, 
                      family = Gamma(link = "log"),
                      control = glmerControl(optimizer = "bobyqa"))

glmer_fit2_8 <- glmer(Burden_inv10 ~ s.lui * s.age + samp + s.nsuccess + (1 | nes), 
                      data = data, 
                      family = Gamma(link = "log"),
                      control = glmerControl(optimizer = "bobyqa"))

glmer_fit2_9 <- glmer(Burden_inv10 ~ s.lui + s.age * samp + s.nsuccess + (1 | nes), 
                      data = data, 
                      family = Gamma(link = "log"),
                      control = glmerControl(optimizer = "bobyqa"))

glmer_fit2_10 <- glmer(Burden_inv10 ~ s.lui + s.age + samp * s.nsuccess + (1 | nes), 
                       data = data, 
                       family = Gamma(link = "log"),
                       control = glmerControl(optimizer = "bobyqa"))

glmer_fit2_11 <- glmer(Burden_inv10 ~ s.lui + s.age + (1 | nes), 
                        data = data, 
                        family = Gamma(link = "log"),
                        control = glmerControl(optimizer = "bobyqa"))

glmer_fit2_12 <- glmer(Burden_inv10 ~ s.lui + s.age + s.nsuccess + (1 | nes), 
                        data = data, 
                        family = Gamma(link = "log"),
                        control = glmerControl(optimizer = "bobyqa"))


## **Check Variance of `nes` for Each Model**

# Create a list of all GLMM models
glmer_models <- list(
  glmer_fit2_null, glmer_fit2_global, glmer_fit2_1, glmer_fit2_2, glmer_fit2_3, 
  glmer_fit2_4, glmer_fit2_5, glmer_fit2_6, glmer_fit2_7, glmer_fit2_8, 
  glmer_fit2_9, glmer_fit2_10, glmer_fit2_11, glmer_fit2_12
)

# Function to check variance of random effect
check_variance <- function(model, model_name) {
  variance <- summary(model)$varcor
  singular <- isSingular(model, tol = 1e-4)
  
  return(data.frame(
    Model = model_name,
    Variance = ifelse(length(variance) > 0, as.numeric(variance[[1]][1]), NA),
    Singular = singular
  ))
}

# Check variance for all models
variance_results <- do.call(rbind, 
                            lapply(seq_along(glmer_models), function(i) {
                              check_variance(glmer_models[[i]], paste0("Model_", i))
                            }))

print(variance_results)

```

Will run a GLM Instead (No Random Effect) Since nes isn't contributing, remove it and use GLM instead of GLMM.

```{r}
# Null model (no random effect)
glm_fit2_null <- glm(Burden_inv10 ~ 1, 
                     data = data, 
                     family = Gamma(link = "log"))

# Global model with all predictors
glm_fit2_global <- glm(Burden_inv10 ~ s.lui + s.age + samp + s.nsuccess, 
                       data = data, 
                       family = Gamma(link = "log"))

# Univariate models
glm_fit2_1 <- glm(Burden_inv10 ~ s.lui, 
                  data = data, 
                  family = Gamma(link = "log"))

glm_fit2_2 <- glm(Burden_inv10 ~ s.age, 
                  data = data, 
                  family = Gamma(link = "log"))

glm_fit2_3 <- glm(Burden_inv10 ~ samp, 
                  data = data, 
                  family = Gamma(link = "log"))

glm_fit2_4 <- glm(Burden_inv10 ~ s.nsuccess, 
                  data = data, 
                  family = Gamma(link = "log"))

# Multivariate models
glm_fit2_5 <- glm(Burden_inv10 ~ s.lui + samp, 
                  data = data, 
                  family = Gamma(link = "log"))

glm_fit2_6 <- glm(Burden_inv10 ~ s.lui + s.nsuccess, 
                  data = data, 
                  family = Gamma(link = "log"))

glm_fit2_7 <- glm(Burden_inv10 ~ samp + s.nsuccess, 
                  data = data, 
                  family = Gamma(link = "log"))

glm_fit2_8 <- glm(Burden_inv10 ~ s.lui * s.age + samp + s.nsuccess, 
                  data = data, 
                  family = Gamma(link = "log"))

glm_fit2_9 <- glm(Burden_inv10 ~ s.lui + s.age * samp + s.nsuccess, 
                  data = data, 
                  family = Gamma(link = "log"))

glm_fit2_10 <- glm(Burden_inv10 ~ s.lui + s.age + samp * s.nsuccess, 
                   data = data, 
                   family = Gamma(link = "log"))

glm_fit2_11 <- glm(Burden_inv10 ~ s.lui + s.age, 
                   data = data, 
                   family = Gamma(link = "log"))

glm_fit2_12 <- glm(Burden_inv10 ~ s.lui + s.age + s.nsuccess, 
                   data = data, 
                   family = Gamma(link = "log"))

```

Use AIC to see the best model Now no singular fit!

```{r}
# Compare GLM model performance using AICc
AIC_table2 <- compare_performance(glm_fit2_null, glm_fit2_global, 
                                  glm_fit2_1, glm_fit2_3, glm_fit2_4, 
                                  glm_fit2_5, glm_fit2_6, glm_fit2_7, 
                                  glm_fit2_8, glm_fit2_9, glm_fit2_10, 
                                  glm_fit2_11, glm_fit2_12, 
                                  rank = TRUE, metrics = "AICc")

# Print the AIC table
print(AIC_table2)

# Save the AIC table as an RDS file
saveRDS(AIC_table2, here("4 Results", "Tables", "ModelAIC2.rds"))

# Convert the AIC table to a formatted flextable
AIC_table2_flextable <- AIC_table2 %>%
  as_flextable() %>%
  set_header_labels(.,
                    Name = "Model",
                    AICc_wt = "AICc Weight",
                    Performance_Score = "Performance Score") %>%
  colformat_num(j = "AICc_wt", digits = 3) %>%  # Format AICc Weight column with 3 decimal places
  colformat_num(j = "Performance_Score", digits = 3) %>%  # Format Performance Score with 3 decimal places
  theme_box() %>%
  autofit()

# Save as a Word document
doc <- read_docx() %>% 
  body_add_par("Table: AICc Model Comparison", style = "heading 1") %>%  # Add a title
  body_add_flextable(AIC_table2_flextable)

# Define the path and save the Word document
print(doc, target = here("4 Results", "Tables", "ModelAIC2.docx"))

# Compare GLM performance and store results
glm_indices <- compare_performance(glm_fit2_null, glm_fit2_global, 
                                   glm_fit2_1, glm_fit2_3, glm_fit2_4, 
                                   glm_fit2_5, glm_fit2_6, glm_fit2_7, 
                                   glm_fit2_8, glm_fit2_9, glm_fit2_10, 
                                   glm_fit2_11, glm_fit2_12)

# Print GLM model comparison results
print(glm_indices)

# Save the performance comparison as a TIFF image
tiff(file = here('4 Results', 'Figures', 'glm_indices.tiff'))
plot(glm_indices)
dev.off()

```

Summarize and graph best model (lowwest AIC) Best fit: 1st: glm_fit2_9 2nd: glm_fit2_global

glm_fit2_9 <- model = glm(Burden_inv10 ~ s.lui + s.age* samp + s.nsuccess, data = data, family = Gamma(link = "log"))

```{r}
# Check GLM model assumptions and diagnostics
glm_check <- check_model(glm_fit2_9)
glm_check

# Save the diagnostic plot
png(file = here('4 Results', 'Figures', 'glm_check.png'))
plot(glm_check)
dev.off()

# Get summary of the GLM model
model2 <- broom::tidy(glm_fit2_9)
print(model2)

# Convert the model summary table to a flextable and format with 3 decimal places
model2_flextable <- model2 %>%
  as_flextable() %>%
  set_header_labels(
    term = "Term",
    estimate = "Estimate",
    std.error = "Standard Error",
    statistic = "t-value"
  ) %>%
  colformat_num(j = c("estimate", "std.error", "statistic"), digits = 3) %>%  # Format numeric columns with 3 decimal places
  theme_box() %>%
  autofit()

# Save as a Word document
doc <- read_docx() %>% 
  body_add_par("Table: Model Summary for glm_fit2_9", style = "heading 1") %>%
  body_add_flextable(model2_flextable)

# Define the path and save the Word document
print(doc, target = here("4 Results", "Tables", "ModelSummary_glm_fit2_9.docx"))

# Save model results
model2 %>% 
  saveRDS(here("4 Results/Models/GLMmodel2.rds"))
```

LUI

```{r}
# Plot model predictions
BLp <- plot_model(glm_fit2_9, terms = 's.lui', type="pred", 
           ci.lvl = 0.95,  # Ensures 95% confidence interval
           show.data = TRUE, show.values = TRUE, show.p = TRUE,
           title = "", show.legend = TRUE, line.size=1,
           axis.title = c("Standarized Landfill Use", 'Sum of Resistance Gene Burden (Inverse Log10)')) + 
           scale_x_continuous(breaks = c(-1, 0, 1, 2),  # Define x-axis breaks based on values
           labels = c("None (-1)", "Mild (0)", "Moderate (1)", "High (2)")) 
BLp

# Save the plot as TIFF using ggsave
ggsave(filename = here('4 Results', 'Figures', 'LUIvsBurden_GLM.tiff'), 
       plot = BLp, device = "tiff", width = 10, height = 6, dpi = 300)
```

Age*Samp

```{r}
# Check and clean the data
summary(data$s.age)  # Ensure no missing or infinite values
data_clean <- data %>% filter(!is.na(s.age) & !is.infinite(s.age))

AgeSamp_plot <- plot_model(glm_fit2_9, terms = c("s.age", "samp"), type = "pred",
           ci.lvl = 0.95,
           show.data = TRUE, show.values = TRUE, show.p = TRUE,
           title = "", show.legend = TRUE, line.size = 1,
           axis.title = c("Standardized Age", "Sum of Resistance Gene Burden (Inverse Log10)")) + 
           
           # Ensure x-axis breaks match previous formatting (as in MLp)
           scale_x_continuous(breaks = c(-2,-1, 0, 1, 2)) + 

           # Adjust y-axis breaks for better clarity
           scale_y_continuous(breaks = seq(0, 3000, by = 500)) +

           # Add jitter to avoid overplotting
           geom_jitter(width = 0.1, height = 0.1, alpha = 0.5) +

           # Add legend title
           labs(color = "Sampling Period")  

AgeSamp_plot

# Save the Age & Sampling Period interaction plot
ggsave(filename = here('4 Results', 'Figures', 'AgeSampvsBurden_GLM.tiff'), 
       plot = AgeSamp_plot, device = "tiff", width = 10, height = 6, dpi = 300)
```

Nest Success

```{r}
# Plot model predictions for Nest Success
NSuccess_plot <- plot_model(glm_fit2_9, terms = 's.nsuccess', type="pred", 
           ci.lvl = 0.95,  
           show.data = TRUE, show.values = TRUE, show.p = TRUE,
           title = "", show.legend = TRUE, line.size=1,
           axis.title = c("Standarized Nest Success", 'Sum of Resistance Gene Burden (Inverse Log10)')) +
           scale_x_continuous(breaks = seq(min(data$s.nsuccess), max(data$s.nsuccess), by=0.5))  
NSuccess_plot

# Save the Nest Success plot
ggsave(filename = here('4 Results', 'Figures', 'NSuccessvsBurden_GLM.tiff'), 
       plot = NSuccess_plot, device = "tiff", width = 10, height = 6, dpi = 300)

```

Odds Ratio Tables for top Models:

glmer_fit_2

odds ratios and their confidence intervals (CI) for glmer_fit_2

```{r}
# Extract coefficients and calculate odds ratios
model1_summary <- broom.mixed::tidy(glmer_fit_2, effects = "fixed")
model1_summary <- model1_summary %>%
  mutate(
    OR = exp(estimate),
    CI_lower = exp(estimate - 1.96 * std.error),
    CI_upper = exp(estimate + 1.96 * std.error)
  )

# Check column names first
colnames(model1_summary)

# Adjust select() if column names differ
model1_OR_table <- model1_summary %>%
  select(any_of(c("term", "estimate", "OR", "CI_lower", "CI_upper", "p.value"))) %>% 
  as_flextable() %>%
  set_header_labels(
    term = "Term",
    estimate = "Estimate (log-OR)",
    OR = "Odds Ratio",
    CI_lower = "CI Lower (95%)",
    CI_upper = "CI Upper (95%)",
    p.value = "P-value"
  ) %>%
  colformat_num(j = c("estimate", "OR", "CI_lower", "CI_upper", "p.value"), digits = 3) %>%
  theme_box() %>%
  autofit()


# Save the OR table as a Word document
doc <- read_docx() %>%
  body_add_par("Table: Odds Ratio Summary for glmer_fit_2", style = "heading 1") %>%
  body_add_flextable(model1_OR_table)

print(doc, target = here("4 Results", "Tables", "OddsRatioTable_glmer_fit_2.docx"))


```

Visualize

```{r}
# Extract coefficients and calculate odds ratios for visualization
model1_summary <- broom.mixed::tidy(glmer_fit_2, effects = "fixed") %>%
  mutate(
    OR = exp(estimate),
    CI_lower = exp(estimate - 1.96 * std.error),
    CI_upper = exp(estimate + 1.96 * std.error)
  )

# Rename terms in model1_summary
model1_summary <- model1_summary %>%
  mutate(term = recode(term, 
                       's.age' = 'AGE'))

# Plot odds ratios with confidence intervals and customized labels
Oddsplot1 <- ggplot(model1_summary, aes(x = reorder(term, OR), y = OR)) +
  geom_point() +
  geom_errorbar(aes(ymin = CI_lower, ymax = CI_upper), width = 0.2) +
  coord_flip() +
  labs(
    title = "Odds Ratios and 95% CI for Top GLMM to predict AMR",
    x = "",
    y = "Odds Ratio (Log Scale)"
  ) +
  scale_y_log10() +  # Use log scale for odds ratios
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5)  # Center the title
  )

# Save the plot as a TIFF file
ggsave(filename = here("4 Results", "Figures", "OddsRatiosPlot.tiff"), 
       plot = Oddsplot1, 
       device = "tiff", 
       width = 10, 
       height = 6, 
       dpi = 300)
```

While generalize linear models (glm_fit2_9) do not directly produce odds ratios like logistic models, you can interpret the coefficients' effects on the response variable:

```{r}
# Extract coefficients for glm_fit2_9 and calculate 95% CI
model2_summary <- broom.mixed::tidy(
    glm_fit2_9, 
    effects = "fixed", 
    conf.int = TRUE, 
    test = TRUE
) %>%
mutate(
    CI_lower = estimate - 1.96 * std.error,
    CI_upper = estimate + 1.96 * std.error,
    p_value = 2 * (1 - pnorm(abs(statistic))) # Calculate two-tailed p-value
)

model2_summary

# Check if p.value exists before selecting it
if("p.value" %in% colnames(model2_summary)) {
  model2_summary_table <- model2_summary %>%
    select(term, estimate, std.error, CI_lower, CI_upper, statistic, p_value) %>%
    as_flextable() %>%
    set_header_labels(
      term = "Term",
      estimate = "Estimate",
      std.error = "Standard Error",
      CI_lower = "95% CI Lower",
      CI_upper = "95% CI Upper",
      statistic = "Z-value",
      p.value = "P-value"
    ) %>%
    colformat_num(j = c("estimate", "std.error", "CI_lower", "CI_upper", "statistic", "p_value"), digits = 3) %>%
    theme_box() %>%
    autofit()
} else {
  model2_summary_table <- model2_summary %>%
    select(term, estimate, std.error, CI_lower, CI_upper, statistic) %>%
    as_flextable() %>%
    set_header_labels(
      term = "Term",
      estimate = "Estimate",
      std.error = "Standard Error",
      CI_lower = "95% CI Lower",
      CI_upper = "95% CI Upper",
      statistic = "Z-value"
    ) %>%
    colformat_num(j = c("estimate", "std.error", "CI_lower", "CI_upper", "statistic"), digits = 3) %>%
    theme_box() %>%
    autofit()
}

# Save the model summary table with 95% CI as a Word document
doc <- read_docx() %>%
  body_add_par("Table: Coefficients Summary for glm_fit2_9 with 95% CI", style = "heading 1") %>%
  body_add_flextable(model2_summary_table)

print(doc, target = here("4 Results", "Tables", "ModelSummary_glm_fit2_9_with_CI.docx"))


```

Visualize glm_fit2_9 <- model: glm(Burden_inv10 ~ s.lui + s.age*samp + s.nsuccess, data = data, family = Gamma(link = "log"))

```{r}
# Extract coefficients for visualization from glm_fit2_9
model2_summary <- broom::tidy(glm_fit2_9, conf.int = TRUE)

# Rename terms and set factor levels to control y-axis order
model2_summary <- model2_summary %>%
  mutate(term = recode(term, 
                       '(Intercept)' = '(Intercept)',
                       's.age' = 'AGE',
                       's.lui' = 'LUI',
                       'samp' = 'SAMPLING PERIOD',
                       's.nsuccess' = 'NEST SUCCESS',
                       's.age:samp' = 'AGE × SAMPLING PERIOD')) %>%
  mutate(term = factor(term, levels = c('LUI', 'SAMPLING PERIOD', 'NEST SUCCESS', 'AGE × SAMPLING PERIOD', '(Intercept)', 'AGE')))  # Specify order

# Plot coefficients with confidence intervals
SumRatiosPlot <- ggplot(model2_summary, aes(x = term, y = estimate)) +
  geom_point() +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2) +  # Uses extracted CIs
  coord_flip() +
  labs(
    title = "Coefficient Estimates and 95% CI for GLM Predicting ARG Burden",
    x = "",
    y = "Standardized Coefficient Estimate"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5)  # Center the title
  )

SumRatiosPlot

# Save the plot as a TIFF file
ggsave(filename = here("4 Results", "Figures", "SumRatiosPlot_GLM.tiff"), 
       plot = SumRatiosPlot, 
       device = "tiff", 
       width = 10, 
       height = 6, 
       dpi = 300)


```
